---
title: "Pumping and Water Table Analyses"
author: "Jill Deines"
date: "April 20, 2018"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

Goal: Analyse the impact of LEMA implementation on well pumping (aggregate) and water table levels.


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path='../figure/01.00_pumpingWtElev_2017/',
                      cache = FALSE)
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(CausalImpact)
library(knitr)

# filepath to repo on local system
mainDir <- '/Users/deinesji/Documents/code_git/1phd/DeinesEtAl2019_ERL_lema1'
```

# Load Data
master data file created in the aptly named 00.40_makeMasterDataFile.Rmd

```{r load}
dataDir <- paste0(mainDir, '/data/tabular')
masterName <- 'masterTidyData_Wide_20181018.csv'

# load
master0 <- read_csv(paste0(dataDir, '/', masterName)) 

# add a column for presence of LEMA program
lemaKey <- data.frame(Year = 1996:2017,
                      isLema = c(rep('before',17),rep('post',5)))
master <- master0 %>% left_join(lemaKey, by = 'Year')

# make a long version of relevant variables
waterUseLong <- master %>%
  dplyr::select(c(masterid, Year, wimasVol_km3, wimasIrrArea_ha, wtElev_m,
                  wtRelative2012_m,wtRelative0_m, pr_calendar_mm, irrigated_ha,
                  aimIrr_ha, annualChange_m)) %>%
  tidyr::gather(., key = variable, value = value, wimasVol_km3:annualChange_m)

# separate regions
sheridan <- master %>% filter(masterid == 'sheridan')
null9 <- master %>% filter(masterid == 'null_geo9')


set.seed(43)
```

# Functions

## causalImpact helper fun
simplify individual runs slightly. note the intervention time is static and specific to LEMA

```{r fun_helper}

runCausalImpact <- function(startYear,endYear, dataframe){
  # make a zoo object for row index thingy
  year <- c(startYear:endYear)
  inputZoo <- zoo(as.matrix(dataframe),  year)

  # run model (intervention year is static)
  before <- c(startYear,2012)
  after <- c(2013,endYear)
  causalModel <- CausalImpact(inputZoo, before, after)
  return(causalModel)
}

```

# Pumping Volumes
Compare aggregate pumping per region by pre-/post-LEMA periods

## Numbers

```{r pumpingChangeTable}
master2002 <- master %>% filter(Year >= 2002)

cropMasterTable <- master2002 %>%
  dplyr::select(c(masterid, isLema, wimasVol_km3, wtElev_m, wtRelative0_m,
                   pr_JanAug_mm, pr_grow_mm, pr_calendar_mm, cornPriceAdj2017,
                   wimasIrrArea_ha, aimIrr_ha, cropActive_ha, aridity,
                  annualChange_m)) %>%
  group_by(masterid, isLema) %>%
  summarise_all(funs(mean), na.rm=TRUE) %>%
  tidyr::gather(., key = variable, value = hectares, 
                wimasVol_km3:annualChange_m) %>%
  tidyr::spread(., key = isLema, value = hectares) %>%
  ungroup() %>%
  group_by(masterid, variable) %>%
  mutate(percDiff = (post - before)/before*100,
         diff = post - before) %>%
  arrange(variable, masterid)

options(scipen=999)    
kable(cropMasterTable, digits = 2)
```


## Causal Impact Analysis
Using the `CausalImpact` package. the first column of the input dataframe needs to be the response variable (here, pumping), and all subsequent columns are taken as covariates

Covariates can't be affected by the intervention

### sheridan pumping
covariates: year, precipitation, aridity, corn prices. 

```{r pumping_sheridan}
# extract df for causalimpact
sheridanPumping <- sheridan %>% 
  filter(Year <= 2017) %>%
  dplyr::select(c(wimasVol_km3, Year, pr_calendar_mm, pr_grow_mm, pr_JanAug_mm,
                   aridity, cornPriceAdj2017)) 

# format data input and run model
sheridanModel <- runCausalImpact(1996, 2017, sheridanPumping)

# see results
plot(sheridanModel) 
summary(sheridanModel)
summary(sheridanModel, 'report')

# what variables were important?
plot(sheridanModel$model$bsts.model, "coefficients")
```

### Null/Control Region pumping
does the intervention have any effect on the null? Using exact same model inputs.

```{r pumping_null}
# extract df for causalimpact
nullPumping <- null9 %>% 
  filter(Year <= 2017) %>%
  dplyr::select(c(wimasVol_km3, Year, pr_calendar_mm, pr_grow_mm, pr_JanAug_mm,
                  aridity, cornPriceAdj2017)) 

# format data input and run model
nullModel <- runCausalImpact(1996, 2017, nullPumping)

# see results
plot(nullModel)  
summary(nullModel)
summary(nullModel, 'report')

# what variables were important?
plot(nullModel$model$bsts.model, "coefficients")
```


# Water Levels

## Causal Impact

### Sheridan Water Table

```{r wt0_sheridan}
# extract df for causalimpact
sheridanwt <- sheridan %>% 
   dplyr::select(c(wtRelative0_m, Year, pr_calendar_mm, pr_grow_mm,
                  pr_JanAug_mm, 
                  aridity, cornPriceAdj2017)) 

# format data input and run model
sheridanModel_wt <- runCausalImpact(1996, 2017, sheridanwt)

plot(sheridanModel_wt) 
summary(sheridanModel_wt)
summary(sheridanModel_wt, 'report')

plot(sheridanModel_wt$model$bsts.model, "coefficients")
```

#### Summary of rate of change
Reviewer Question: did the rate of decline change?

```{r sheridanMeanChange}
# get annual changes
sheridanwt_lag <- sheridanwt %>%
  mutate(wtlag = wtRelative0_m - lag(wtRelative0_m))

# summarize for pre-lema ref period
sheridanwt_lag %>%
  filter(Year <= 2012 & Year > 2002) %>%
  summarize(mean(wtlag))

# don't include 2013 since that is the change from 2012 = 2012 pre-lema pumping
sheridanwt_lag %>%
  filter(Year <= 2017 & Year > 2013) %>%
  summarize(mean(wtlag))
```


### Null Water Table

```{r wt0_null}
# extract df for causalimpact
nullwt <- null9 %>% 
  dplyr::select(c(wtRelative0_m,Year,  pr_calendar_mm, pr_grow_mm,
                  pr_JanAug_mm, 
                  aridity, cornPriceAdj2017))

# format data input and run model
nullModel_wt <- runCausalImpact(1996, 2017, nullwt)


plot(nullModel_wt) 
summary(nullModel_wt)
summary(nullModel_wt, 'report')
```

#### Summary of rate of change
Reviewer Question: did the rate of decline change?

```{r nullMeanChange}
# get annual changes
nullwt_lag <- nullwt %>%
  mutate(wtlag = wtRelative0_m - lag(wtRelative0_m))

# summarize for pre-lema ref period
nullwt_lag %>%
  filter(Year <= 2012 & Year > 2002) %>%
  summarize(mean(wtlag))

# don't include 2013 since that ist he change from 2012 = 2012 pre-lema pumping
nullwt_lag %>%
  filter(Year <= 2017 & Year > 2013) %>%
  summarize(mean(wtlag))
```

# Combine Model Output

## Extract tables

### function to pull data

```{r fun_modelextract}
extractModelNums <- function(causalModel1, variable1, masterid1, startYear){
  data1 <- as.data.frame(causalModel1$series[,c('response','point.pred',
                                 'point.pred.lower', 'point.pred.upper')])
  data1$variable <- variable1
  data1$masterid <- masterid1
  data1$year <- startYear:(startYear+nrow(data1)-1)
  return(data1)
}
```

### extract and merge

```{r pumping}
# get model output
sherPump <- extractModelNums(sheridanModel, 'pumping','LEMA', 1996)
nullPump <- extractModelNums(nullModel, 'pumping','Control', 1996)
sherWT <- extractModelNums(sheridanModel_wt, 'wtElev','LEMA', 1996)
nullWT <- extractModelNums(nullModel_wt, 'wtElev','Control', 1996)

# combine and export
allModels <- rbind(sherPump,nullPump, sherWT, nullWT)

write_csv(allModels, paste0(dataDir,'/causalImpact_summaries_05-20_20181018.csv'))
```



