---
title: "Area and Crop Changes"
author: "Jill Deines"
date: "April 23, 2018"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

Goal: Summarize changes in irrigated area and crop choices before/after LEMA


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path='../figure/02.00_areaCropChoices_2017/',
                      cache = FALSE)
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
library(lubridate)
library(knitr)

# filepath to repo on local system
mainDir <- '/Users/deinesji/Documents/code_git/1phd/DeinesEtAl2019_ERL_lema1'
```

# Load Data
master data file created in the aptly named 00.40_makeMasterDataFile.Rmd

```{r load}
dataDir <- paste0(mainDir, '/data/tabular')
masterName <- 'masterTidyData_Wide_20181018.csv'

# load
master0 <- read_csv(paste0(dataDir, '/', masterName)) 

# add a column for presence of LEMA program
lemaKey <- data.frame(Year = 1996:2017,
                      isLema = c(rep('before',17),rep('post',5)))
master <- master0 %>%
  left_join(lemaKey, by = 'Year')


# subset for 5 year before/after
master5 <- master %>% filter(Year > 2007)
sheridan5 <- master5 %>% filter(masterid == 'sheridan') %>% arrange(Year)
control5 <- master5 %>% filter(masterid == 'null_geo9') %>% arrange(Year)
```

# Crops

## Master Table
Display before, after, and % change for pre and post means

```{r cropMasterTable}
cropMasterTable <- master5 %>%
  dplyr::select(-c(wimasVol_km3, wtElev_m, wtRelative0_m, wtRelative2012_m,
                   pr_JanAug_mm, pr_grow_mm,
                   totalArea_ha, nonAg_ha, Year, Alfalfa_irrigated_ha,
                   Alfalfa_rainfed_ha, annualChange_m)) %>%
  group_by(masterid, isLema) %>%
  summarise_all(funs(mean), na.rm=TRUE) %>%
  tidyr::gather(., key = variable, value = hectares, 
                wimasIrrArea_ha:cornPriceAdj2017) %>%
  tidyr::spread(., key = isLema, value = hectares) %>%
  ungroup() %>%
  group_by(masterid, variable) %>%
  mutate(percDiff = (post - before)/before*100,
         diff = post - before) %>%
  arrange(variable, masterid)

options(scipen=999)    
kable(cropMasterTable, digits = 2)
```

Calculate corn's percentage of irrigated area pre/post LEMA in Sheridan

```{r cornyChanges}
cropMasterTable %>%
  ungroup() %>%
  filter(masterid == 'sheridan' & variable %in% 
           c('aimIrr_ha', 'Corn_irrigated_ha', 'wimasIrrArea_ha'))

# irr area mean pre-lema
irrPre <- mean(8985,10058)
irrPost <- mean(8825,9589)

# corn proportion preLema
6892/irrPre

# corn proportion postLema
6005/irrPost
```


# Irrigated Area

Examine changes in irrigated area. Compare estimates from WIMAS and AIM-RRB.

## 5 years before (2008-2017)

Summarize WIMAS

```{r irrWimas}
# summary
master5 %>%
  group_by(masterid, isLema)  %>%
  summarize(Mean = mean(wimasIrrArea_ha, na.rm=TRUE),
            q25 = quantile(wimasIrrArea_ha, .25, na.rm=TRUE),
            q50 = quantile(wimasIrrArea_ha, .5, na.rm=TRUE),
            q75 = quantile(wimasIrrArea_ha, .75, na.rm=TRUE),
            min = min(wimasIrrArea_ha, na.rm=TRUE),
            max = max(wimasIrrArea_ha, na.rm=TRUE))

# rough stats
wimIrr <- master5 %>%
  filter(Year <= 2017) %>%
  group_by(masterid)  %>%
  do(tidy(t.test(wimasIrrArea_ha~isLema, data =.)))
wimIrr
```

Summarise AIM

```{r irrAim}
master5 %>%
   group_by(masterid, isLema)  %>%
  summarize(Mean = mean(aimIrr_ha, na.rm=TRUE),
            q25 = quantile(aimIrr_ha, .25, na.rm=TRUE),
            q50 = quantile(aimIrr_ha, .5, na.rm=TRUE),
            q75 = quantile(aimIrr_ha, .75, na.rm=TRUE),
            min = min(aimIrr_ha, na.rm=TRUE),
            max = max(aimIrr_ha, na.rm=TRUE))

# rough stats
wimIrr <- master %>%
  filter(Year %in% 2008:2017) %>%
  group_by(masterid)  %>%
  do(tidy(t.test(aimIrr_ha~isLema, data =.)))
wimIrr
```


plot them

```{r irrArea_2008}

# get mean irrigated area, before and after
meanIrrArea <- master %>%
  filter(Year > 2007) %>%
  dplyr::select(masterid, isLema, wimasIrrArea_ha, aimIrr_ha) %>%
  gather(., key = dataset, value = value, wimasIrrArea_ha:aimIrr_ha)

# plot wimas
ggplot(meanIrrArea %>% filter(dataset == 'wimasIrrArea_ha'), 
       aes(x = isLema, y = value)) +
  geom_boxplot() + 
  facet_wrap(~masterid) +
  theme_bw() + 
  ggtitle('WIMAS Change in Area, 2008-2012 vs 2013-2016')

# plot aim
ggplot(meanIrrArea %>% filter(dataset == 'aimIrr_ha'), 
       aes(x = isLema, y = value)) +
  geom_boxplot() + 
  facet_wrap(~masterid) +
  theme_bw() + 
  ggtitle('AIM-RRB Change in Area, 2008-2012 vs 2013-2017')

# plot both
ggplot(meanIrrArea, 
       aes(x = isLema, y = value, group = interaction(dataset, isLema), color = dataset)) +
  geom_boxplot() + 
  facet_wrap(~masterid) +
  theme_bw() + 
  ggtitle('Change in Area by Dataset, 2008-2012 vs 2013-2017')
```


