---
title: "Water per crop - irrigation depth analysis"
author: "Jill Deines"
date: "April 26, 2018"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

Goal: Check out changes in irrigation water applied by crop via causalImpact analysis (to control for annual climate)


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path='../figure/03.00_cropWater_2017/',
                      cache = FALSE)
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(CausalImpact)

# filepath to repo on local system
mainDir <- '/Users/deinesji/Documents/code_git/1phd/DeinesEtAl2019_ERL_lema1'
```

# Load
uses formatted wimas data from 00.35_wells_WIMAS.Rmd (PDIV_ID well id column removed to preserve anonymity) and gets precip from master data file from 00.40_makeMasterDataFile.Rmd

```{r load}
# wimas ---------------------------------------------------
wellDir <- paste0(mainDir, '/data/wellData/WIMAS_AnthonyCleaned_JillFormatted')

# load cleaned well data 
wells<- read_csv(paste0(wellDir,
                  '/WIMAS_1996-2017_SheridanNull9_convertedUnits_noPDIV_ID.csv'))

# add a column for presence of LEMA program
lemaKey <- data.frame(year = 1996:2017,
                      isLema = c(rep('before',17),rep('post',5)))
wellsPeriod <- wells %>%
  left_join(lemaKey, by = 'year')

# remove NA's
wellsComplete <- wellsPeriod[!is.na(wellsPeriod$volume_m3),]

# climate covariables ------------------------------------------
dataDir <- paste0(mainDir, '/data/tabular')
masterName <- 'masterTidyData_Wide_20181018.csv'

# load
master <- read_csv(paste0(dataDir, '/', masterName)) %>%
  dplyr::select(c(masterid,Year,pr_calendar_mm, pr_JanAug_mm, 
                  pr_grow_mm, aridity, cornPriceAdj2017)) 
  

```

# Functions

## causalImpact helper fun
simplify individual runs slightly. note the intervention time is static and specific to LEMA

```{r fun_helper}

runCausalImpact <- function(startYear,endYear, dataframe){
  set.seed(5)
  # make a zoo object for row index thingy
  year <- c(startYear:endYear)
  inputZoo <- zoo(as.matrix(dataframe),  year)

  # run model (intervention year is static)
  before <- c(startYear,2012)
  after <- c(2013,endYear)
  causalModel <- CausalImpact(inputZoo, before, after)
  return(causalModel)
}

```


# Corn

## Get Mean Depth
Calculate average irrigation depth for corn fields based on WIMAS data.
In WIMAS, fields that are 100% corn are '2'

Other codes of interest:

* 3 = grain sorghum
* 4 = soybeans
* 5 = wheat


## Prepare for CausalImpact
add climate variables and split into regions

```{r mergeClimate}
# mean depth by region
cornDepth <- wellsComplete %>%
  filter(crop == '2' & depth_m >0) %>%
  group_by(masterid, year) %>%
  summarise(meanDepth_cm = mean(depth_m*100))

startYear <- 1996
endYear <- 2017
cornVars <- cornDepth %>%
  filter(year > startYear-1) %>%
  rename(Year = year) %>%
  left_join(master, by = c('masterid','Year'))

sheridanCorn <- cornVars %>% 
  ungroup() %>%
  filter(masterid == 'sheridan') %>%
  dplyr::select(-c(masterid, Year))  %>%
  mutate(year = startYear:endYear)
controlCorn <- cornVars %>% 
  ungroup() %>%
  filter(masterid == 'null_geo9') %>%
  dplyr::select(-c(masterid, Year)) 
```

## Causal: Sheridan
Question: Did LEMA intervention affect LEMA corn depth?

```{r causal_lemaDepth}
# format data input and run model
LemaCornModel <- runCausalImpact(startYear, endYear, sheridanCorn)

# see results
plot(LemaCornModel)  
summary(LemaCornModel)
#summary(LemaCornModel, 'report')

# what variables were important?
#plot(LemaCornModel$model$bsts.model, "coefficients")
```

## Causal: Control
Question: Did LEMA intervention affect control corn depth?

```{r causal_controlDepth}
# format data input and run model
controlCornModel <- runCausalImpact(1996, 2017, controlCorn)

# see results
plot(controlCornModel)  
summary(controlCornModel)
#summary(controlCornModel, 'report')

# what variables were important?
#plot(controlCornModel$model$bsts.model, "coefficients")
```

# Soybeans

## Get Mean Depth

```{r soyDepths}
# mean depth by region
soydepth <- wellsComplete %>%
  filter(crop == '4' & depth_m >0) %>%
  group_by(masterid, year) %>%
  summarise(meanDepth_cm = mean(depth_m*100))

startYear <- 1996
endYear <- 2017
soyVars <- soydepth %>%
  filter(year > startYear-1) %>%
  rename(Year = year) %>%
  right_join(master, by = c('masterid','Year'))

sheridanSoy <- soyVars %>% 
  ungroup() %>%
  filter(masterid == 'sheridan' &
           Year <= endYear) %>%
  dplyr::select(-c(masterid, Year))  %>%
  mutate(year = startYear:endYear)
controlSoy <- soyVars %>% 
  ungroup() %>%
  filter(masterid == 'null_geo9' &
           Year <= endYear) %>%
  dplyr::select(-c(masterid, Year))  %>%
  mutate(year = startYear:endYear)
```

## Causal soy
Question: Did LEMA intervention affect LEMA soy depth?

```{r causal_lemaDepth_soy}
# format data input and run model
LemaSoyModel <- runCausalImpact(startYear, endYear, sheridanSoy)

# see results
plot(LemaSoyModel)  
summary(LemaSoyModel)
#summary(LemaSoyModel, 'report')

# what variables were important?
#plot(LemaSoyModel$model$bsts.model, "coefficients")
```

## Causal: Control
Question: Did LEMA intervention affect control soy depth?

```{r causal_controlDepth_soy}
# format data input and run model
controlSoyModel <- runCausalImpact(startYear, endYear, controlSoy)

# see results
plot(controlSoyModel)  
summary(controlSoyModel)
#summary(controlSoyModel, 'report')

# what variables were important?
#plot(controlCornModel$model$bsts.model, "coefficients")
```


# Wheat

## Get Mean Depth

```{r wheatDepths}
# mean depth by region
wheatdepth <- wellsComplete %>%
  filter(crop == '5' & depth_m >0) %>%
  group_by(masterid, year) %>%
  summarise(meanDepth_cm = mean(depth_m*100))

startYear <- 1996
endYear <- 2017
wheatVars <- wheatdepth %>%
  filter(year > startYear-1) %>%
  rename(Year = year) %>%
  right_join(master, by = c('masterid','Year'))

sheridanWheat <- wheatVars %>% 
  ungroup() %>%
  filter(masterid == 'sheridan' &
           Year <= endYear) %>%
  dplyr::select(-c(masterid, Year)) %>%
  mutate(year = startYear:endYear)
controlWheat <- wheatVars %>% 
  ungroup() %>%
  filter(masterid == 'null_geo9' &
                      Year <= endYear) %>%
  dplyr::select(-c(masterid, Year)) %>%
  mutate(year = startYear:endYear)
```

## Causal Wheat
Question: Did LEMA intervention affect LEMA wheat depth?

```{r causal_lemaDepth_wheat}
# format data input and run model
LemaWheatModel <- runCausalImpact(startYear, 2016, sheridanWheat)

# see results
plot(LemaWheatModel)  
summary(LemaWheatModel)
#summary(LemaWheatModel, 'report')

# what variables were important?
#plot(LemaWheatModel$model$bsts.model, "coefficients")
```

## Causal: Control
Question: Did LEMA intervention affect control wheat depth?

```{r causal_controlDepth_wheat}
# format data input and run model
controlWheatModel <- runCausalImpact(startYear, endYear, controlWheat)

# see results
plot(controlWheatModel)  
summary(controlWheatModel)
#summary(controlWheatModel, 'report')

# what variables were important?
#plot(controlWheatModel$model$bsts.model, "coefficients")
```

# Sorghum
See what this looks like with minimal pre-lema data

## Get Mean Depth

```{r sorghumDepths}
# mean depth by region
sorghumdepth <- wellsComplete %>%
  filter(crop == '3' & depth_m >0) %>%
  group_by(masterid, year) %>%
  summarise(meanDepth_cm = mean(depth_m*100))

startYear <- 1996
endYear <- 2017
sorghumVars <- sorghumdepth %>%
  filter(year > startYear-1) %>%
  rename(Year = year) %>%
  right_join(master, by = c('masterid','Year'))

sheridanSorghum <- sorghumVars %>% 
  ungroup() %>%
  filter(masterid == 'sheridan' &
           Year <= endYear) %>%
  dplyr::select(-c(masterid, Year)) %>%
  mutate(year = startYear:endYear)
controlSorghum <- sorghumVars %>% 
  ungroup() %>%
  filter(masterid == 'null_geo9' &
                      Year <= endYear) %>%
  dplyr::select(-c(masterid, Year)) %>%
  mutate(year = startYear:endYear)
```

## Causal Sorghum
Question: Did LEMA intervention affect LEMA wheat depth?

```{r causal_lemaDepth_sorghum}
# format data input and run model
LemaSorghumModel <- runCausalImpact(startYear, endYear, sheridanSorghum)

# see results
plot(LemaSorghumModel)  
summary(LemaSorghumModel)
#summary(LemaSorghumModel, 'report')

# what variables were important?
#plot(LemaSorghumModel$model$bsts.model, "coefficients")
```

## Causal: Control
Question: Did LEMA intervention affect control sorghum depth?

```{r causal_controlDepth_sorghum}
# format data input and run model
controlSorghumModel <- runCausalImpact(startYear, 2016, controlSorghum)

# see results
plot(controlSorghumModel)  
summary(controlSorghumModel)
#summary(controlSorghumModel, 'report')

# what variables were important?
#plot(controlSorghumModel$model$bsts.model, "coefficients")
```


