---
title: 'Wizard Wells: Kriging'
author: "Jill Deines"
date: "April 17, 2018"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

Goal: Krige water levels from the KS Wizard Well database

Update 5/27/2018: re-run with re-processed well dataset.

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path='../figure/03.15_actualKriging/',
                      cache = FALSE)
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(rgdal)
library(gstat)
library(latticeExtra)
library(raster)
```

# Load Data

## Well Data and AOI

```{r loadData}
# use the HPA-CLASS project projection
classProj <- '+proj=aea +lat_1=20 +lat_2=60 +lat_0=40 +lon_0=-101 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0'

# buffer boundary
gisDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMA_Part1/data/GIS/boundaries'
AOIs.ll <- readOGR(gisDir,'Combined_Null9_S6_10kmBuff_indvlPolys')
AOIs <- spTransform(AOIs.ll, CRS(classProj))

# filtered wizard wells
wellDir <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMA_Part1/data/wellData/WIZARD/cleaned'
wells.ll <- readOGR(wellDir, 'wells_cleaned_clipped_1996-2017_jimRemoved')
wells <- spTransform(wells.ll, CRS(classProj))

# add projected coordinates to df
wells.coords <- coordinates(wells)
wells$Xloc <- wells.coords[,'coords.x1']
wells$Yloc <- wells.coords[,'coords.x2']

plot(AOIs)
plot(wells, add=TRUE)
```


# Functions

## Make Interpolation Grids

This uses Ashton Shortridge's custom function for making a SpatialPixelsDataFrame from a polygon:

```{r FUN_buildGrid}

# build.SpinPolyGrid
# A function that constructs a SpatialPixelsDataFrame for a grid of points
# within the boundaries of a (multi) SpatialPolygons object.
# poly is the SpatialPolygons object, gridres is the spatial resolution of each
# pixel, and coordNames is a vector of names for the resulting coordinates.
# Returns a SpatialPixelsDataFrame.
build.SpinPolyGrid <- function(poly, gridres=1000, coordNames = c('x', 'y')) {
    grid.df <- makegrid(poly, cellsize=gridres)
    names(grid.df) <- coordNames
    grid.df$id <- 1:length(grid.df[,1])  # an ID column.
    coordinates(grid.df) <- coordNames # convert to a SpatialPoints Object
    grid.df@proj4string <- poly@proj4string # If there's coord system info, set that.
    class(poly) <- "SpatialPolygons"
    grid.pts <- grid.df[!is.na(over(grid.df, poly)),]
    return(as(grid.pts, "SpatialPixelsDataFrame")) # Cast to SpatialPixelsDataFrame
}

```

# Krige!
Methods optimized in 03.10_wells_wizardKriging.Rmd or whatever it was called

Do this in a loop for all years available (2000-2017)

## variables

```{r set_vars}
# file location and names
outDirGis <- 'C:/Users/deinesji/Dropbox/1PhdJill/hpa/LEMA_Part1/data/GIS/waterLevels'
fSuffixElev <- '_waterTableElev_wizardKrige_m_jimRemoved.tif'
fSuffixChange <- '_waterTableChange_wizardKrige_m_jimRemoved'

# settings
resolution <- 250 # in m
yearStart <- 1996
yearEnd <- 2017
```

## Universal Kriging
Kriging is based on universal kriging with a simple model incorporting easting to account for the large anisotropy due to the E-W trend in elevation

```{r uk, warning=FALSE, cache=TRUE}
years <- yearStart:yearEnd
for (year in years){
  # subset wells
  wellsub <- wells[wells$YearAdj == year,]

  # make year specific trend surface model
  tsm <- lm(wtElev_m ~ Xloc, data = wellsub)
 
  # make grid for predictions
  grid250 <- build.SpinPolyGrid(AOIs, gridres = resolution)
  grid250$Xloc <- coordinates(grid250)[,'x']

  # fit variogram on residuals
  vgram <- variogram(resid(tsm)~1, 
                     locations = coordinates(wellsub),
                     data = wellsub)
  vgramModel <- fit.variogram(vgram, vgm(c('Sph', 'Gau','Exp')))
  

  print(plot(vgram, vgramModel, 
       main = paste0(year, ' Vgram Model:', vgramModel[2,1],
                    ' \n Nug=', round(vgramModel[1,2],2), ' psill=',
                    round(vgramModel[2,2],2), ' range=', 
                    round(vgramModel[2,3],0))))

  # universal kriging, simple model ------------------
  simpleKrig <- krige(wtElev_m ~ Xloc, wellsub, grid250, model = vgramModel)
  
  # plot hack
  predictions <- spplot(simpleKrig, "var1.pred",
                        main=paste(year, " UK wtElev ~ X")) +
    layer(sp.polygons(AOIs)) + layer(sp.points(wellsub, col='black'))
  
  var <- spplot(simpleKrig, "var1.var", main="variance")+
    layer(sp.polygons(AOIs)) + layer(sp.points(wellsub))

  plot(predictions, split = c(1,1,2,1), more=T)
  plot(var, split=c(2,1,2,1), more=F)
  
  # convert output to raster and export
  ras <- raster(simpleKrig[,'var1.pred'])
  writeRaster(ras, paste0(outDirGis,'/',year,fSuffixElev))
  
  # clean up in case variogram fails?
  rm(vgramModel)
}
```
