---
title: "WIMAS Water Use"
author: "Jill Deines"
date: "April 16, 2018"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

Goal: Subset WIMAS to study area and do some formatting.

update 10-18-2018: re-running with wimas data through 2017; and filtering by source. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path='../../figure/00.35_wimas_2017/',
                      cache = FALSE)
library(knitr)
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse) # for ggplot2, tidyr, dplyr
library(sf)
library(lubridate)
library(tmap)

sessionInfo()
```

# Load Data
Formatted WIMAS data (Anthony + Jill pre-cleaning) and AOI shapefile produced in 03.01_wells_WIZARD. Formatted data produced by:

* downloaded the full WIMAS database from http://hercules.kgs.ku.edu/geohydro/wimas/query_setup.cfm, annually
* formatting with 00.30_wells_WIMAS_preprocessing.py (Anthony Kendall)
* reshaping with 00.33_wells_WIMAS_preprocessing2.Rmd (Jill Deines)

## Subset WIMAS Data
Subset WIMAS for study area only to create a smaller file to be used in subsequent analyses. Run once.

Also exports lists of wells in study regions

```{r loadData, eval=FALSE}
# directories
projFolder <- '/Users/deinesji/Documents/code_git/1phd/DeinesEtAl2019_ERL_lema1'
gisDir <- paste0(projFolder, '/data/GIS/boundaries')
wellDir <- paste0(projFolder, '/data/wellData/WIMAS_AnthonyCleaned_JillFormatted')
proj <- 4267 
startYear <- 1996
endYear <- 2017

# well data (orriginally in 4267, NAD27)
wimas0 <- read_csv(paste0(wellDir, '/WIMAS_1990-2017_gwPts_ksHpa_withSource.csv')) %>%
  st_as_sf(coords = c('LONGITUDE','LATITUDE'), crs = proj)
#st_write(wimas0, paste0(wellDir, '/WIMAS_1990-2017_gwPts_ksHpa_withSource.shp'))
wimas <- wimas0 %>% filter(source == 'GW')

# boundaries of interest (orriginally in 4269, NAD83)
AOIs <- read_sf(paste0(gisDir,'/Combined_Null9_S6_10kmBuff_indvlPolys.shp')) %>%
  st_transform(proj)

# extract points in both aois
wellsIn <- st_join(wimas, AOIs) %>% filter(!is.na(masterid))

# keep data from startYear on
wellsInTime <- wellsIn %>%
  filter(year >= startYear)

# unique wells by region?
wellsInTime %>% 
  group_by(masterid) %>%
  summarize(n = n_distinct(PDIV_ID))

# export culled data as shapefile, csv, 
write_sf(wellsInTime, paste0(wellDir,'/WIMAS_', startYear,'-', endYear, '_Lema1Buffer.shp'))
write_sf(wellsInTime, paste0(wellDir,'/WIMAS_', startYear,'-', endYear, '_Lema1Buffer.csv'))

# and list of well IDs

# get unique pdiv_ids: full buffer
wellIds <- data.frame(PDIV_ID = unique(wellsInTime$PDIV_ID))
write.csv(wellIds, 
          paste0(wellDir,'/wellList/wimas_wellIDs_buffer_',startYear,'_on_', endYear,'.csv'), 
          row.names=FALSE)

# get unique pdiv_ids: lema and null9
wellsInTime.df <- wellsInTime
st_geometry(wellsInTime.df) <- NULL
regionWellIds <- wellsInTime.df %>%
  filter(masterid %in% c('null_geo9','sheridan')) %>%
  distinct(PDIV_ID)
write.csv(regionWellIds, paste0(wellDir,'/wellList/wimas_wellIDs_lemaNull_',
                              startYear, '_',endYear,'.csv'), 
          row.names=FALSE)
```

## Format a bit more
Make a file with just Sheridan and the Null9 region; will further tabulate in the 05.10_makeMasterDataFile.Rmd

Also convert units, add crop key

```{r formatData2}
# directories
gisDir <- '/Users/deinesji/Dropbox/1PhdJill/hpa/LEMA_Part1/data/GIS/boundaries'
wellDir <- '/Users/deinesji/Dropbox/1PhdJill/hpa/LEMA_Part1/data/wellData/WIMAS_AnthonyCleaned_JillFormatted'
proj <- 4267 
startYear <- 1996
endYear <- 2017

# well data (orriginally in 4267, NAD27)
wimas <- read_sf(paste0(wellDir, '/WIMAS_1996-2017_Lema1Buffer.shp'))

# boundaries of interest (orriginally in 4269, NAD83)
AOIs <- read_sf(paste0(gisDir,'/Combined_Null9_S6_10kmBuff_indvlPolys.shp')) %>%
  st_transform(proj)

# plot for funzies
tm_shape(AOIs) + tm_polygons('masterid') + tm_shape(wimas) + tm_dots(size=.3)

# remove buffer wells
wimas2 <- wimas %>% filter(masterid != 'buffer')

# convert units
wimas3 <- wimas2 %>% 
  mutate(area_m2 = acres * 4046.86,
         depth_m = depth * 0.3048,
         volume_m3 = volume * 1233.48) %>%
  select(c(masterid, year, PDIV_ID, crop, volume_m3, area_m2, depth_m, system))

# # # export this for future use
# write_sf(wimas3, paste0(wellDir,'/WIMAS_', startYear,
#                         '-',endYear,'_SheridanNull9_convertedUnits.shp'))
# write_sf(wimas3, paste0(wellDir,'/WIMAS_', startYear,
#                         '-',endYear,'_SheridanNull9_convertedUnits.csv'))

```

## Count wells by region

```{r WellCount}
yearCounts <- wimas3 %>%
  as.data.frame() %>%
  group_by(masterid, year) %>%
  summarize(wells = n())

kable(yearCounts)
```

### count wells with data
more specifically, count wells that have data at least 1x in study period

```{r wellCount2}
# filter out wells without data during study period
wellsWithData <- wimas3 %>%
  ungroup() %>%
  group_by(PDIV_ID) %>%
  # count years of NA's
  summarise(missingYears = sum(is.na(volume_m3))) 

# check out missing year counts
table(wellsWithData$missingYears)

# get list of wells with no data in all 21 years from 1996-2017
wellsWithData2 <- wellsWithData %>%  filter(missingYears < 21)



yearCounts2 <- wimas3 %>%
  as.data.frame() %>%
  # remove wells without data during 1996-2017
  filter(PDIV_ID %in% wellsWithData2$PDIV_ID) %>%
  group_by(masterid, year) %>%
  summarize(wells = n())

kable(yearCounts2)
```

