---
title: 'Data Sauron: Make Master Data File'
author: "Jill Deines"
date: "April 19, 2018"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

Goal: Combine data sets into a master file for analysis

Update 10/18/2018: re-done with WIMAS data including 2017

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path='../../figure/00.40_makeMasterData_2017/',
                      cache = FALSE)
```

**R Packages Needed**

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)

# filepath to repo on local system
mainDir <- '/Users/deinesji/Documents/code_git/1phd/DeinesEtAl2019_ERL_lema1'

sessionInfo()
```

# Get to It

## CDL
area by CDL type and irrigation status (aim-rrb) was exported using GEE script 00.04_gee_CDL_Summaries_final.js

### Load and format

2015 and 2015 have extra cdl class columns: 251,252,253. In all national
metadata files for 2008 - 2016, these values are all NA. So, dropping columns.
most come with a 0 column which is nonsense, so removing so all match.
2013 has a class 62 that is NA for all others, for 'pasture/grass', but no
classified cells for this class, so removing that as well.

```{r processCDL}
# directories
dataDir <- paste0(mainDir, '/data/tabular')
# cdl key 
cdlkey <- read.csv(paste0(dataDir, '/CDL_key_2014.csv'))

# CDL  annual files -  read in to a single dataframe
cdlfiles <- list.files(paste0(dataDir,'/cdl_tables_lema'),
            pattern="*Sheridan_Null9_area_irrigationStatus_tractsOnly_giFixed",
                       full.names = TRUE)

# read in to a single dataframe, removing junk columns and first dummy row
dataAll <- as.data.frame(do.call(rbind, lapply(cdlfiles, function(x) {
  csv1 <- read.csv(x)
  csv1 <- csv1[, -which(names(csv1)%in% c('system.index','.geo','X0','X62',
                                          'X251','X252','X253'))]
  csv1 <- csv1[-1,]    # remove dummy row
  csv2 <- csv1[csv1$masterid %in% c('sheridan','null_geo9'),]
})))  

# convert to long data format and remove NA's
dataLong <- gather(dataAll, key = cdlCode, value = area_m2, X1:X92)
dataLong <- dataLong[!is.na(dataLong$area_m2),]

# add a crop type column using cdl key
dataLong$cdlCode <- as.numeric(substr(dataLong$cdlCode, start = 2, 
                                      stop = nchar(dataLong$cdlCode)))
cdlkey$cdlClassName <- as.character(cdlkey$CLASS_NAME)
dataLong2 <- merge(dataLong, cdlkey[,c('VALUE','cdlClassName')],
                   by.x = 'cdlCode', by.y = 'VALUE')

# reogranize
cdlLong <- dataLong2 %>%
  select(c(masterid, Year, status, cdlCode, cdlClassName, area_m2))
```

### Derive CDL totals
Factors of interest:

* annual active cropland
* annual fallow
* annual irrigated
* annual rainfed
* irrigated/nonirrigated corn
* irrigted/nonirrigated sorghum
* irrigated/nonirrigated soy
* irrigated/nonirrigated wheat
* irrigated/nonirrigated alfalfa

```{r cdlAggregating}
# included switchgrass (60) in non-ag, grass pasture (176)
nonAg <- c(0,60,63,64,65, 81, 82, 83, 87, 88, 92, 111, 112, 121,122,123,124,131,
           141, 142, 143, 152, 176, 190, 195)

fallow <- 61
doubleCrop <- c(225,226, 230:241, 254)

top5 <- c(1,4,5,24,36) # corn, sorghum, soy, wheat, alfalfa

# summarize categories (this feels inefficient but im uninspired)
nonAgTotal <- cdlLong %>%
  group_by(masterid, Year) %>%
  filter(cdlCode %in% nonAg) %>%
  summarize(nonAg_ha = sum(area_m2*1e-4))

fallowtotal <- cdlLong %>%
  group_by(masterid, Year) %>%
  filter(cdlCode %in% fallow) %>%
  summarize(fallow_ha = sum(area_m2*1e-4))

doubleCropTotal <- cdlLong %>%
  group_by(masterid, Year) %>%
  filter(cdlCode %in% doubleCrop) %>%
  summarize(doubleCropped_ha = sum(area_m2*1e-4))

activeAg <- cdlLong %>%
  group_by(masterid, Year) %>%
  filter(!(cdlCode %in% c(nonAg,fallow))) %>%
  summarize(cropActive_ha = sum(area_m2*1e-4))

# get ag by irrigated status and start for format for combining
agByStatus <- cdlLong %>%
  filter(!(cdlCode %in% c(nonAg,fallow))) %>%
  group_by(masterid, Year, status) %>%
  summarize(area_ha = sum(area_m2*1e-4)) %>%
  spread(., key = status, value = area_ha) %>%
  rename(irrigated_ha = irrigated, rainfed_ha = rainfed)

# get top 5 crops by irrigated status and start for format for combining
top5IrrStatus <- cdlLong %>%
  filter(cdlCode %in% top5) %>%
  group_by(masterid, Year, status, cdlClassName) %>%
  summarize(cropArea_ha = sum(area_m2*1e-4)) %>%
  mutate(crop_status = 
           gsub(" ", "", paste(cdlClassName,status,'ha',sep = '_'))) %>%
  ungroup() %>%
  select(-c(status, cdlClassName)) %>%
  spread(., key = crop_status, value = cropArea_ha)

# add 'em together: wide format
cdlVars <- cdlLong %>%
  group_by(masterid, Year) %>%
  summarize(totalArea_ha = sum(area_m2*1e-4)) %>%
  left_join(activeAg, by = c('masterid','Year')) %>%
  left_join(fallowtotal, by = c('masterid','Year')) %>%
  left_join(doubleCropTotal, by = c('masterid','Year')) %>%
  left_join(nonAgTotal, by = c('masterid','Year')) %>%
  left_join(agByStatus, by = c('masterid','Year')) %>%
  left_join(top5IrrStatus, by = c('masterid','Year'))

# question: how much of irrigated area is represented by top 4 crops?
means <- cdlVars %>%
  ungroup() %>%
  filter(Year >= 2008) %>%
  group_by(masterid) %>%
  summarize(meanActiveCrop = mean(cropActive_ha),
            meanIrrCrop = mean(irrigated_ha),
            meanCornirr = mean(Corn_irrigated_ha),
            meanSorgIrr = mean(Sorghum_irrigated_ha),
            meanWheatIrr = mean(WinterWheat_irrigated_ha),
            meanSoyIrr = mean(Soybeans_irrigated_ha),
            meanAlfIrr = mean(Alfalfa_irrigated_ha))

means %>%
  rowwise() %>%
  mutate(main4Area = sum(meanCornirr, meanSorgIrr, meanWheatIrr, meanSoyIrr),
         percentOfIrr = main4Area/meanIrrCrop*100)

# alfalfa percent of irrigated area
means$meanAlfIrr/means$meanIrrCrop
```

## Other Data
Add water levels, ancillary variables like precip and aridity, commodity prices(?) aim_rrb

### AIM derived climate and area
Summaries of irrigated area from AIM-RRB and associated ancillary variables extracted in GEE script 00.07_gee_Get_AIMRRB_andAncillary_Final.js; the ancillary variables sampled  here are from the script 00.06_gee_makeAncillaryData, which was used to generate climate summaries for the full High Plains Aquifer + additional regions for a related project. 


```{r AIMcovars, message=FALSE}
#   annual files -  read in to a single dataframe
aimFiles <- list.files(paste0(dataDir,'/aimrrb_plusAncillary_lema'),
                       pattern="*TractMaskedGiFixed")

# read in to a single dataframe, removing junk columns and first dummy row
aimAll <- as.data.frame(do.call(rbind, lapply(aimFiles, function(x) {
  csv1 <- read_csv(paste0(dataDir,'/aimrrb_plusAncillary_lema/',x))
  # remove junk and "non irrigated" column (since cdlVar has non-irr cropland)
  csv2 <- csv1[, -which(names(csv1)%in% c('system:index','.geo', 'groups', '0'))]
  # don't include data for buffer zone
  csv3 <- csv2[csv2$masterid %in% c('sheridan','null_geo9'),]
  # files prior to 1999 (aim-rrb start year) are missing a `1` irrigated column
  if(!(1 %in% names(csv3))) {
    csv3$`1` <- NA
  }
  # rename irrigated
  csv3$aimIrr_ha <- csv3$`1` *1e-4
  # add a year column based on file name
  csv3$Year <- as.numeric(substr(x, start=1,stop=4))
  # arrange output
  csv4 <- csv3 %>%
    select(c(masterid, Year, aimIrr_ha, pr_calendar, pr_ann, pr_grow, aridity))
})))  

# combine with cdlVars
oneData00 <- aimAll %>%
  left_join(cdlVars, by = c('masterid','Year'))
```

### Commodity Prices
2017 inflation Adjusted corn prices (as a proxy for all commodity prices, due to high correlation) derived from NASS stats in 00.10_getCornPrices.Rmd

```{r prices}
# load and format
prices <- read_csv(paste0(dataDir,'/commodityPrices/NASS_corn_prices_1996-2017_inflationAdjusted_2017dollars.csv')) %>%
  mutate(cornPriceAdj2017 = Value) %>%
  select(c(Year,cornPriceAdj2017))

# combine (should fill in rows for sheridan and null)
oneData0 <- oneData00 %>%
  left_join(prices, by = 'Year')
```


### Water Levels
Processed, kriged, and summarized in scripts 00.2X_wells_wizardXXXXX.Rmd.

```{r addMoreData}
# load water levels
waterLevels <- read_csv(paste0(dataDir,'/waterLevels_1996-2017_wizardKriged_lemaNull9_jimRemoved.csv'))

# add to massive data frame
oneData000 <- waterLevels %>%
  left_join(oneData0, by = c('masterid','Year')) %>%
  rename(pr_calendar_mm = pr_calendar,
         pr_JanAug_mm = pr_ann,
         pr_grow_mm = pr_grow)


```

### WIMAS and Pumping
Processed and cleaned in scripts 00.3X_wells_WIMAS_XXXX.Rmd

Get summed volume and area by region. Note raw data is not provided but can be obtained at the link provided in 00.30_wells_WIMAS_preprocessing.Rmd

```{r wimas}
wellDir <- paste0(mainDir, '/data/wellData/WIMAS_AnthonyCleaned_JillFormatted')

# load and sum
wimas <- read_csv(paste0(wellDir,
                         '/WIMAS_1996-2017_SheridanNull9_convertedUnits.csv')) %>%
  rename(Year = year) %>%
  group_by(masterid, Year) %>%
  summarize(wimasVol_km3 = sum(volume_m3 * 1e-6, na.rm = TRUE),
            wimasIrrArea_ha = sum(area_m2 * 1e-4, na.rm=TRUE))

# combine
oneData <-  wimas %>%
  right_join(oneData000, by = c('masterid','Year')) %>%
  arrange(Year, masterid)

oneData

names(oneData)
```

Bonus: analysis changes in irrigation tech for reviewer

```{r irrTech}
wellDir <- paste0(mainDir, '/data/wellData/WIMAS_AnthonyCleaned_JillFormatted')

systemKey <- data.frame(system=1:9,
                        systemCode = c('flood','trickledrip','center pivot','center pivot lepa',
                                       'sprinkler other than cp','center pivot and flood',
                                       'subsurface drip in combo with other','other',
                                       'center pivot with mobile drip'))

# load 
wimas2 <- read_csv(paste0(wellDir,
                         '/WIMAS_1996-2017_SheridanNull9_convertedUnits.csv')) %>%
  left_join(systemKey) %>%
  #filter(year >= 2008) %>%
  filter(system != -99)

# get area by system
systemArea <- wimas2 %>%
  group_by(masterid, year, system) %>%
  summarize(area_km2 = sum(area_m2)*1e-6) %>%
  filter(area_km2 >0) 

# plot
ggplot(systemArea,
       aes(x = year, y = area_km2, color = as.factor(system), group = as.factor(system))) +
  geom_line() +
  facet_wrap(~masterid) +
  theme_bw()

# compare percent increase in lepa
# add a column for presence of LEMA program
lemaKey <- data.frame(year = 1996:2017,
                      isLema = c(rep('before',17),rep('post',5)))
lepa <- systemArea %>%
  left_join(lemaKey) %>%
  filter(system ==4 & year >= 2008)

ggplot(lepa,
       aes(x = year, y = area_km2)) +
  geom_line() +
  facet_wrap(~masterid) +
  theme_bw()

nullLepa <- lm(area_km2~year, data = lepa[lepa$masterid == 'null_geo9',])
summary(nullLepa)

sd6Lepa <- lm(area_km2~year, data = lepa[lepa$masterid == 'sheridan',])
summary(sd6Lepa)


```

## Export
This df will be used for subsequent analyses

```{r export}
write_csv(oneData, paste0(dataDir, '/masterTidyData_Wide_aimTractMaskedGiFixed_jimRemoved_with2017_20181018.csv'))
```

